---
title: "Replicating payoff of digital option"
description: |
  We explore payoff replication of exotic options using a portfolio on vanilla options. We investigate how well the replication is and what factors affect the payoff.
author:
  - name: Sherman Ding
    url: https://www.linkedin.com/in/shermanding93/
    affiliation: Derivative Pricing
    affiliation_url: https://www.cfainstitute.org/en/membership/professional-development/refresher-readings/basics-derivative-pricing-valuation#:~:text=Derivatives%20are%20priced%20by%20creating,eliminates%20any%20possibility%20of%20arbitrage.
date: 01-03-2022
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

\newcommand{\ci}{\perp\!\!\!\perp}

```{r}
library(reticulate)
```

# 1.0 Define digital option payoff

We consider a digital option with a payoff at maturity T = 1, where L = 80, U = 120. L and U and lower barriers and upper barriers respectively. The digital option pays $1 if Stock price remains in between the barriers.

\[
1_{L < S_T < U}
\]

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
import matplotlib.pyplot as plt
import numpy as np

def binary_call_payoff(S_T):
  if S_T > 80 and S_T < 120:
    return 1.0
  else:
    return 0.0
  
price = []
payoff = []

for stock in np.arange(70,130,0.1):
    price.append(stock)
    payoff.append(binary_call_payoff(stock))
    
plt.plot(price,payoff)
plt.title('Digital Option Payoff')
plt.show()
```

# 2.0 Replicate payoff of digital option

We replicate the digital option payoff above by combining european vanilla option spreads:

* Long vertical call spread (long call @ $80 - \epsilon$, short call @ $80 + \epsilon$)
* Short vertical call spread (short call @ $80 - \epsilon$, long call @ $80 + \epsilon$)

\[
C_0(Digital) = \lim_{\epsilon \rightarrow 0} \frac{Call(80 - \epsilon) - Call(80 + \epsilon)}{2\epsilon} = -\frac{dCall}{dK}
\]

The digital call can be thought of as a limit of a call spread. As the distance between the call spread strikes and the digital strikes $\epsilon$ get smaller, $\frac{1}{\epsilon}$ call spreads with 2$\epsilon$ width are needed to replicate the digital option. This means in the limit, as $\epsilon$ tends to 0, the call spread can replicate the digital option exactly.
Below plot shows the replication of a digital option payoff using 2 call spreads; as $\epsilon$ becomes smaller, the spreads replicate the digital option payoff more and more perfectly.

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
payoff_list1 = []
payoff_list2 = []
payoff_list3 = []
s_price1 = []
s_price2 = []
s_price3 = []
epsilon_var=(0.1,0.5,1)

for j in range(len(epsilon_var)):
    S=70
    s_tick=0.1
    payout = 0.5 # per call spread leg (total payout=1)
    epsilon=epsilon_var[j]
    lots = payout*2/epsilon 
    K1=80-epsilon
    K2=80+epsilon
    K3=120-epsilon
    K4=120+epsilon
    
    for i in np.arange(70.0,130,s_tick):
        payoff=0
        S+=s_tick
        if S>K1:
            payoff+=(payout*lots)*(S-K1)
        if S>K2:
            payoff-=(payout*lots)*(S-K2)
        if S>K3:
            payoff-=(payout*lots)*(S-K3)
        if S>K4:
            payoff+=(payout*lots)*(S-K4)
        if epsilon == 0.1:      
            payoff_list1.append(payoff)
            s_price1.append(S)
        elif epsilon == 0.5:
            payoff_list2.append(payoff)
            s_price2.append(S)
        elif epsilon == 1:
            payoff_list3.append(payoff)
            s_price3.append(S)
    
plt.plot(s_price1,payoff_list1)
plt.plot(s_price2,payoff_list2)
plt.plot(s_price3,payoff_list3)
plt.title('Spread Replication Payoff')
plt.legend(['epsilon = 0.1','epsilon = 0.5','epsilon = 1'])
plt.show()
```


It is important to note that while $\epsilon$ is not 0, it does not perfectly replicate a digital option payoff:

* Above the barrier level, the call spread has the same payoff as the digital option
* Below the barrier level, the call spread the same payoff as the digital option
* In-between the upper and lower strikes, the call spread has a non-zero payoff

The call-spread over-replicates the digital option because its payoff is always greater, or equal to the digital option payoff.

# 3.0 Black-Scholes Delta and Vega profiles of digital option

It is shown that the smaller the $\epsilon$, the closer the call spread is in replicating the digital option payoff.
By using $\epsilon$ = 0.1, we can set the call spread strikes required for the digital option.

* K1 = 80 - 0.1 = 79.9
* K2 = 80 + 0.1 = 80.1
* K3 = 80 - 0.1 = 119.9
* K4 = 80 + 0.1 = 120.1

Since $\epsilon$ is set at 0.1, we would need 1/$\epsilon$ amount of contracts per call spread.
Therefore we have 1/0.1 = **10 contracts per call spread**

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
from scipy.stats import norm
import numpy as np
from matplotlib import style

sigma = 0.2
r = 0
q = 0
T = 1
s_tick = 0.1
K1 = 79.9
K2 = 80.1
K3 = 119.9
K4 = 120.1
delta_profile = []
stock_price = []

def BS_Call_Delta(F,K,q,sigma,T):
    d1 = (np.log(F/K)+(0.5*sigma**2*T)) / sigma*np.sqrt(T)
    return np.exp(-q*T) * norm.cdf(d1)

for S in np.arange(0.1,240,s_tick):
    Forward = S * np.exp(-r*T)
    call_spread1 = 10*(BS_Call_Delta(Forward,K1,q,sigma,T) - BS_Call_Delta(Forward,K2,q,sigma,T))
    call_spread2 = 10*(-BS_Call_Delta(Forward,K3,q,sigma,T) + BS_Call_Delta(Forward,K4,q,sigma,T))
    delta_profile.append(call_spread1 + call_spread2)
    stock_price.append(S)

with plt.style.context('seaborn'):    
    plt.plot(stock_price,delta_profile)
    plt.title('Digital Option Delta Profile (BS)')
plt.show()
```

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
sigma = 0.2
r = 0
q = 0
T = 1
s_tick = 0.1
K1 = 79.9
K2 = 80.1
K3 = 119.9
K4 = 120.1
vega_profile = []
stock_price_ = []

def BS_Call_Vega(S0,F,K,q,sigma,T):
    d1 = (np.log(F/K)+(0.5*sigma**2*T)) / sigma*np.sqrt(T)
    return np.exp(-q*T) * S0 * np.sqrt(T) * norm.pdf(d1)

for S in np.arange(0.1,240,s_tick):
    Forward = S * np.exp(-r*T)
    call_spread1 = 10*(BS_Call_Vega(S,Forward,K1,q,sigma,T) - BS_Call_Vega(S,Forward,K2,q,sigma,T))
    call_spread2 = 10*(-BS_Call_Vega(S,Forward,K3,q,sigma,T) + BS_Call_Vega(S,Forward,K4,q,sigma,T))
    vega_profile.append(call_spread1 + call_spread2)
    stock_price_.append(S)

with plt.style.context('seaborn'):    
    plt.plot(stock_price_,vega_profile)
    plt.title('Digital Option Vega Profile (BS)')
plt.show()
```