---
title: "Static replication using Breeden-Litzenberger"
description: |
  We use Breeden-Litzenberger to compute option prices numerically and replicate payoffs statically.
author:
  - name: Sherman Ding
    url: https://www.linkedin.com/in/shermanding93/
    affiliation: Derivative Pricing
    affiliation_url: https://www.cfainstitute.org/en/membership/professional-development/refresher-readings/basics-derivative-pricing-valuation#:~:text=Derivatives%20are%20priced%20by%20creating,eliminates%20any%20possibility%20of%20arbitrage.
date: 03-20-2022
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

\newcommand{\ci}{\perp\!\!\!\perp}

```{r}
library(reticulate)
```

# 1.0 Breeden-Litzenberger formula

BL formula states that, any payoff that depends on $S_T$ can be priced with the below formula:

\[
V_0 = e^{-rT} V_T \left(\left(F_0(T)\right) \right) + \int^{F_0(T)}_{0} Put(K,T)\frac{\delta^2V_T(K)}{\delta K^2} dK + \int^{\infty}_{F_0(T)} Call(K,T)\frac{\delta^2V_T(K)}{\delta K^2} dK
\]

In the examples below, we assume:

1. r = q = 0

2. $S_0$ = 1

3. Implied Volatility $\sum{(K)} = 0.510 - 0.591K + 0.376K^2 - 0.105K^3$

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
import numpy as np
from scipy.stats import norm
from scipy.integrate import quad

def ivol_calc(K):
    if K >3:
        return 0.510 - 0.591*3 + 0.376*3**2 - 0.105*3**3 + 0.011*3**4
    else:
        return 0.510 - 0.591*K + 0.376*K**2 - 0.105*K**3 + 0.011*K**4
    
def Black76Call(F, K, T, DF):
    ivol = ivol_calc(K)
    d1 = (np.log(F/K)+(0.5*ivol**2)*T) / (ivol*np.sqrt(T))
    d2 = d1 - ivol*np.sqrt(T)
    return DF*(F*norm.cdf(d1) - K*norm.cdf(d2))
          
def Black76Put(F, K, T, DF):
    ivol = ivol_calc(K)
    d1 = (np.log(F/K)+(0.5*ivol**2)*T) / (ivol*np.sqrt(T))
    d2 = d1 - ivol*np.sqrt(T)
    return DF*(K*norm.cdf(-d2) - F*norm.cdf(-d1)) 
```

## 2.0 Static replication for option with payoff: $V_T(S_T) = \sqrt(S_T)$


$V_T(S_T) = \sqrt{S_T}$

The upper limit of a call integral can be approximated by $F_0(T)e^{\kappa\sigma\sqrt{T}}$
where $\kappa$ represents the number of standard deviation away from the mean, and $\sigma$ is ATM implied vol. The higher $\kappa$ is, the more accurate the approximation is.

Using Breeden-Litzenberger formula, we can use static replication for the payoff $V_T(S_T) = \sqrt{S_T}$ below:

\[
V_0 = \sqrt{S_0}e^{\frac{-rT}{2}} + \int^{F_0(T)}_{0} -\frac{1}{4K^{3/2}} Put(K,T) dK + \int^{\infty}_{F_0(T)} -\frac{1}{4K^{3/2}} Call(K,T) dK
\]

$h(S_T) = S_T^n$

$h'(S_T) = nS_T^{n-1}$

$h''(S_T) = n(n-1)S_T^{n-2}$

$h(F) = S_0^ne^{nrT}$

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
def h_1(ST):
    return np.sqrt(ST) * np.exp(-r*T/2)
          
def Numerical_integration1(S0, r, q, T, SD):
    DF = np.exp(-r*T)
    DivF = np.exp(-q*T)
    F = S0 * DivF/DF
    upper_limit = F * np.exp(SD * ivol_calc(F) * np.sqrt(T))
    putintegrand = lambda x: -x**(-1.5) * Black76Put(F, x, T, DF) / 4
    I_put, error = quad(putintegrand, 0, F)
    callintegrand = lambda y: -y**(-1.5) * Black76Call(F, y, T, DF) / 4
    I_call, error = quad(callintegrand, F, upper_limit)
    return h_1(F) + I_put + I_call

S0 = 1
r = 0
q = 0
T = 4

SDs = np.linspace(1, 6, 6)
print([Numerical_integration1(S0, r, q, T, SD) for SD in SDs])
print('Option price: %.8f' % [Numerical_integration1(S0, r, q, T, SD) for SD in SDs][-1])
```
We see that as kappa increases, the approximation becomes more accurate.

## 3.0 Static replication for option with payoff: $V_T(S_T) = S_T^3$

Using Breeden-Litzenberger formula, we can use static replication for the payoff $V_T(S_T) = S_T^3$ below:

\[
V_0 = S_0^3e^{2rT} + \int^{F_0(T)}_{0} 6K Put(K,T) dK + \int^{\infty}_{F_0(T)} 6K Call(K,T) dK
\]

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
def h_2(ST):
    return ST**3 * np.exp(2*r*T)
          
def Numerical_integration2(S0, r, q, T, SD):
    DF = np.exp(-r*T)
    DivF = np.exp(-q*T)
    F = S0 * DivF/DF
    upper_limit = F * np.exp(ivol_calc(F) * SD * np.sqrt(T))
    putintegrand = lambda x: 6 * x * Black76Put(F, x, T, DF)
    I_put, error = quad(putintegrand, 0, F)
    callintegrand = lambda y: 6 * y * Black76Call(F, y, T, DF)
    I_call, error = quad(callintegrand, F, upper_limit)
    return h_2(F) + I_put + I_call

S0 = 1
r = 0
q = 0
T = 4

SDs = np.linspace(1, 6, 6)
print([Numerical_integration2(S0, r, q, T, SD) for SD in SDs])
print('Option price: %.8f' % [Numerical_integration2(S0, r, q, T, SD) for SD in SDs][-1])
```
Again, we see that as kappa increases, the approximation becomes more accurate.