---
title: "Modelling probability of an event using poisson process"
description: |
  We look at how to model the probabability of an event by using jumps in a poisson process.
author:
  - name: Sherman Ding
    url: https://www.linkedin.com/in/shermanding93/
date: 2022-07-18
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\newcommand{\ci}{\perp\!\!\!\perp}

```{r}
library(reticulate)
```

# 1.0 The Poisson Process

Assuming the event to be exactly one asset jump arrival within the interval $\left[0,T\right]$, how can we determine the probability of this jump:

We consider a Poisson process and we denote the time of the n-th jump by $T_n$. The sequence {$T_1,T_2,...T_n$} is the sequence of arrival(or jump) times. The number of events in a time interval $\Delta t$ is poisson distributed with mean $\lambda \Delta t$ for all t $\ge$ 0, with arrival intensity $\lambda$:

\[
Prob\left(N_{t+\Delta t} - N_t = n\right) = e^{-\lambda \Delta \frac{T}{2}} \frac{\left(\lambda \Delta \frac{T}{2}\right)^n}{n!}, n = 0,1,...
\]

**Event A**
Let Event A be the probability of 1 asset jump in a given time interval $\left[0,T\right]$, n = 1.

\[
Prob\left(N_{t+\Delta t} - N_t = 1\right) = e^{-\lambda T} \frac{\left(\lambda T\right)^1}{1!}
\\
Prob\left(A\right) = \lambda Te^{-\lambda T}
\]

The probability of Event A is $\lambda Te^{-\lambda T}$.

# 2.0 Properties of the poisson process, and mathematical proof

How can we prove that the probability of the event of exactly one asset jump within the interval $\left[0,T\right]$ is the same as:

1. 1 asset jump in the interval $\left[0,\frac{T}{2}\right)$, and no asset jump in the interval $\left[\frac{T}{2},T\right]$.

2. No asset jumps in the interval $\left[0,\frac{T}{2}\right)$, and 1 asset jump in the interval $\left[\frac{T}{2},T\right]$.

We consider a Poisson process and we denote the time of the n-th jump by $T_n$. The sequence {$T_1,T_2,...T_n$} is the sequence of arrival(or jump) times. The number of events in a time interval $\Delta t$ is poisson distributed with mean $\lambda \Delta t$ for all t $\ge$ 0:

\[
Prob\left(N_{t+\Delta t} - N_t = n\right) = e^{-\lambda \Delta \frac{T}{2}} \frac{\left(\lambda \Delta \frac{T}{2}\right)^n}{n!}, n = 0,1,...
\]

Expectation (Mean) is given as: E$\left[N_{t+\Delta t} - N_t\right] = \lambda \Delta t$

We know that the Poisson process is:

* Independent and Identically distributed
* A discrete probability distribution of a sum of independent Bernoilli trials

This means that at any point in time, the Poisson process probabilistically restarts itself.

**Event X**

Let Event X be the probability of exactly 1 jump from interval $\left[0,\frac{T}{2}\right)$ and 0 jump from interval $\left[\frac{T}{2},T\right]$, where $X_1$ = 1 and $X_2$ = 0:

First, we find the probability of 1 jump occuring in the interval $\left[0,\frac{T}{2}\right)$:

\[
Prob\left(N_t = 1\right) = Prob\left(X_1 = 1\right)= e^{-\lambda \frac{T}{2}}\frac{\left(\lambda \frac{T}{2}\right)^n}{n!} = e^{-\lambda \frac{T}{2}} \frac{\lambda T}{2}
\]

For interval $\left[\frac{T}{2},T\right]$, the interval length is the same as the first interval $\left[0,\frac{T}{2}\right]$. We can find the probability of 0 jump occuring in this interval:

\[
Prob\left(N_t = 0\right) = Prob\left(X_2 = 0\right) = e^{-\lambda \frac{T}{2}}
\]

We can find the probability of **Event X** occuring below:

Where **Event X** is Prob($X_2 = 0$) conditional on Prob($X_1 = 1$) for interval  $\left[0,\frac{T}{2}\right)$

\[
Prob\left(X_1 = 1\right) \times Prob\left(X_2 = 0\right) = e^{-\lambda \frac{T}{2}} \frac{\lambda T}{2} \times e^{-\lambda \frac{T}{2}} 
\\
\\
Prob\left(X\right) = Prob\left(X_2 = 0 | X_1 = 1\right) = e^{-\lambda T} \left(\frac{\lambda T}{2}\right)
\]

<br>

**Event Y**

Let Event Y be the probability of 0 jump from interval $\left[0,\frac{T}{2}\right)$ and exactly 1 jump from interval $\left[\frac{T}{2},T\right]$:

First, we find the probability of 0 jump occuring in the interval $\left[0,\frac{T}{2}\right)$:

\[
Prob\left(N_t = 0\right) = Prob\left(Y_1 = 0\right) = e^{-\lambda \frac{T}{2}}
\]

For interval $\left[\frac{T}{2},T\right]$, the interval length is the same as the first interval $\left[0,\frac{T}{2}\right]$. We can find the probability of 1 jump occuring in this interval:

\[
Prob\left(N_t = 1\right) = Prob\left(Y_2 = 1\right) = e^{-\lambda \frac{T}{2}}\frac{\left(\lambda \frac{T}{2}\right)^n}{n!} = e^{-\lambda \frac{T}{2}} \frac{\lambda T}{2}
\]

We can find the probability of **Event Y** occuring below:

Where Prob($Y_2 = 1$) is conditional on Prob($Y_1 = 0$) for interval  $\left[0,\frac{T}{2}\right)$

\[
Prob\left(Y_1 = 0\right) \times Prob\left(Y_2 = 1\right) = e^{-\lambda \frac{T}{2}} \times e^{-\lambda \frac{T}{2}} \frac{\lambda T}{2}
\\
\\
Prob\left(Y\right) = Prob\left(Y_2 = 1 | Y_1 = 0\right) = e^{-\lambda T} \left(\frac{\lambda T}{2}\right)
\]

### Probability Comparison

Event A:

$Prob\left(A\right) = \lambda Te^{-\lambda T}$

Event X or Event Y:

$Prob\left(X \cup Y\right) = Prob\left(X\right) + Prob\left(Y\right) = e^{-\lambda T} \left(\frac{\lambda T}{2}\right) + e^{-\lambda T} \left(\frac{\lambda T}{2}\right) = \lambda Te^{-\lambda T}$

# 3.0 Programming the mathematical proof

We can assign arbitrary numbers to the variables and conduct below checks:

Set $\lambda$ = 0.5, T = 1, $\mu$ = $\lambda \frac{T}{2}$

* Check that probabilities of both Events X and Y are the same
* Calculate probabilities and check against inbuilt Poisson PMF function from Scipy Library

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
import numpy as np

lamb = 0.5
T = 1
mu = lamb*(T/2)

# Event X probabilities
Prob_X1 = np.exp(-lamb*T/2)*(((lamb*T/2)**1)/1)
Prob_X2 = np.exp(-lamb*T/2)

# Event Y probabilities
Prob_Y1 = np.exp(-lamb*T/2)
Prob_Y2 = np.exp(-lamb*T/2)*(((lamb*T/2)**1)/1)

print('Probability of 1 asset jump in interval [0,T/2]: '+str(Prob_X1))
print('Probability of 0 asset jump in interval [T/2,0]: '+str(Prob_X2))
print('Probability of Event X: '+str(Prob_X1*Prob_X2))
print('Probability of 1 asset jump in interval [0,T/2]: '+str(Prob_Y1))
print('Probability of 0 asset jump in interval [T/2,0]: '+str(Prob_Y2))
print('Probability of Event Y: '+str(Prob_Y1*Prob_Y2))
```

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
# Part A probability
Prob_A = np.exp(-lamb*T)*(lamb*T)

print('Probability of Event A: '+str(Prob_A))
print('Probability of Event X or Event Y: '+str(Prob_X1*Prob_X2 + Prob_Y1*Prob_Y2))
```

### We can use the Poisson Probability Mass Function from the Scipy library to calculate the individual event probability for each aportioned intervals:

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
from scipy.stats import poisson

# Event X
PMF_X1 = poisson.pmf(1, mu)
PMF_X2 = poisson.pmf(0, mu)

# Event Y
PMF_Y1 = poisson.pmf(0, mu)
PMF_Y2 = poisson.pmf(1, mu)

print('Probability of 1 asset jump in interval [0,T/2]: '+str(PMF_X1))
print('Probability of 0 asset jump in interval [T/2,0]: '+str(PMF_X2))
print('Probability of Event X: '+str(PMF_X1*PMF_X2))
print('Probability of 0 asset jump in interval [0,T/2]: '+str(PMF_Y1))
print('Probability of 1 asset jump in interval [T/2,0]: '+str(PMF_Y2))
print('Probability of Event Y: '+str(PMF_Y1*PMF_Y2))
```

### Finally:

```{python, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
# Event A vs (Event X or Event Y)
PMF_A = poisson.pmf(1, mu*2)
print('Probability of Event A: '+str(PMF_A))
print('Probability of Event X or Event Y in interval [0,T]: '+str(PMF_X1*PMF_X2 + PMF_Y1*PMF_Y2))
```
Hence, we have proved that the event of 1 asset jump in interval $\left[0,T\right]$ in Part A gives rise to the exact same probability as:

* The event of 1 jump in the interval $\left[0,\frac{T}{2}\right)$ and no jumps in interval $\left[\frac{T}{2},T\right]$

**or**

* The event of 0 asset jump in interval $\left[0,\frac{T}{2}\right)$ and 1 jump in interval $\left[\frac{T}{2},T\right]$ 
